---
title: "A two-dimensional integrand with a diagonal ridge"
description: |
  This document looks the performance of the serial CUHRE and VEGAS algorithms
  on a function that has a diagnoal ridge.
author:
  - name: Marc Paterno
    url: https://github.com/marcpaterno
date: 12-15-2020
updated: 12-16-2020
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(rmarkdown)
library(tidyverse)
```


## Purpose of this analysis

The serial VEGAS algorithm typically is challenged by integrands that are not
separable, and which contain ridges that are not aligned with the coordinate
axes. This analysis uses such an integrand to demonstrate this issue.

The integrand chosen is:
$$\frac{4 k u^2}{\frac{1}{100}+\left(u-v -\frac{1}{3}\right)^2}$$
with
$$k = 4050 / \left(43200+14920 \pi -125840 \tan^{-1}(\frac{3}{70})-14920 \tan^{-1}(\frac{3}{50})+110920 \tan^{-1}(\frac{3}{10})+ \\
54000 \tan^{-1}(\frac{50}{3})+54000 \tan^{-1}(\frac{70}{3})+17946 \log(109)-3573 \log(2509)-14373 \log(4909) \right).$$
The approximate value of $k$ is 0.01890022674239546529975841.

The volume of integration is (-1, 1) for both $x$ and $y$.
The correct value of the integral is 1.

## Testing environment

These tests were run on a MacBook Pro laptop.

```{r child = '../../mac130389-description.Rmd'}
```

## Description of the dataframe

```{r child = '../../dataframe-description.Rmd'}
```

```{r read_data, echo = FALSE, message = FALSE}
d <- read_tsv("h2_vegas_cuhre.tsv.xz") %>%
  mutate(r = errorest/(epsrel*value),
         alg = factor(alg, levels =c("vegas", "cuhre_0", "cuhre_1")),
         converged = (r < 1)) %>%
  arrange(desc(epsrel), alg)

paged_table(d)
```

## Analysis

As one measure of the innate "efficiency" of each algorithm,
we consider how many function evaluations are required
to obtain a given fractional error tolerance.

```{r, echo = FALSE}
ggplot(d, aes(1/epsrel, neval, color = alg)) +
  geom_point() +
  scale_x_log10() +
  scale_y_log10() +
  labs(x = "Inverse of fractional error tolerance", y = "Number of function evaluations required")
```

For this integrand, except at the very loosest tolerances
VEGAS always requires more function calls than does CUHRE.
CUHRE with flag=0 requires fewer function evaluations than it does with flag=4.

### Calculation time as a function of required error tolerance

```{r, echo = FALSE}
ggplot(d, aes(1/epsrel, time, color=alg)) +
  geom_point() +
  scale_x_log10() +
  scale_y_log10() +
  labs(x="Inverse of error tolerance", y="Calculation time (ms)")

```

### Error estimate and function evaluations

```{r echo = FALSE, message = FALSE}
d %>%
ggplot(aes(neval, errorest, color = alg)) +
  geom_point() +
  scale_x_log10() +
  scale_y_log10() +
  labs(x = "Number of integrand evaluations", "Fractional error estimate")
```

### Reliability of error estimate

If we look only at the calculations that converged, how accurate are the error
estimates?

```{r, echo=FALSE}
filter(d, !is.na(value)) %>%
  ggplot(aes(1/epsrel, abs(error)/errorest, color = alg)) +
  scale_y_log10(labels = scales::comma) +
  scale_x_log10() +
  geom_point() +
  labs(x = "Inverse of error tolerance",
       y = "Ratio of actual error to estimated error")
```

If the error estimate is reliable, then the ratio should not exceed 1.
For this integrand, it appears that all error estimates are reliable.
In fact, they appear to be *over* estimates.

### Error estimate as a function of running time

Users do not really care about the number of function evaluations done;
they care about the time the algorithm takes to converge for a given
error tolerance, and that the algorithm has actually converged to that tolerance.

```{r, echo = FALSE}
filter(d, !is.na(value)) %>%
  ggplot(aes(1/epsrel, time, color = alg)) +
  scale_x_log10() +
  scale_y_log10(labels = scales::comma) +
  geom_point() +
  labs(x = "Inverse of error tolerance",
       y = "Calculation time (ms)")
```

For this integrand, the CUHRE algorithm is much faster than VEGAS.
There seems to be little performance difference between CUHRE with flag = 0
and flag = 4.

### Regions used by the CUHRE algorithm

It is remarkable how few regions the CUHRE algorithm needs to evaluate this
integral with high accuracy. 


```{r}
filter(d, !is.na(value), alg != "vegas") %>%
  ggplot(aes(1/epsrel, nregions, color = alg)) +
  scale_x_log10() +
  scale_y_log10(labels = scales::comma) +
  geom_point() +
  labs(x = "Inverse of error tolerance",
       y = "Number of regions")
```

