---
title: "Comparing the new CUBA release to the old"
description: |
  In Novemer of 2020, release CUBA 4.2.1 was released. This article compares the
  performance of CUBA 4.2.1 to that of CUBA 4.2.
author:
  - name: Marc Paterno
    url: https://github.com/marcpaterno
date: 12-22-2020
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(rmarkdown)
library(tidyverse)
```

## Purpose of this article

In November of 2020, CUBA version 4.2.1 was released.
This version included a number of "minor fixes", and a correction of some of
the tables of coefficients used by the CUHRE algorithm.
In this article we compare the behavior of the newest version of CUBA to data
previously collec`ted with CUBA 4.2.

## Integrands and volumes of integration

The integrands used for this article are:

* fun6 : $k (u \, v + (x \, y \, w^y)/(1+u) +z^2)$, integrated over the unit volume
* genz_1abs_5d : $| \cos(4 v +5 w + 6 x +7 y + 8 z)|$, integrated over the unit volume
* h2 : $(4 k u^2)/(\frac{1}{100}+\left(u-v -\frac{1}{3}\right)^2)$, integrated over $-1<x<1$, $-1<y<1$.

These are the same integrands used in our earlier articles.

## Read the data

```{r}
read_both <- function(igrand) {
  x0 <- vroom::vroom(sprintf("%s_cuba-4.2.tsv.xz", igrand), comment = "#")
  x0$version = "4.2"
  x1 <- vroom::vroom(sprintf("%s_cuba-4.2.1.tsv.xz", igrand), comment = "#")
  x1$version = "4.2.1"
  x <- bind_rows(x0, x1)
  x$integrand = igrand
  x
}
```

```{r}
x <- bind_rows(lapply(c("fun6", "genz1abs_5d", "h2"), read_both)) %>%
  mutate(integrand = as_factor(integrand),
         alg = as_factor(alg),
         version = as.factor(version),
         neval = as.integer(neval),
         nregions = as.integer(nregions)) %>%
  select(integrand, alg, version, epsrel, value, errorest, error, neval, nregions, time)

paged_table(x)
```

## Comparison of error estimates

The quality of the error estimate for each algorithm is of the highest import,
so we consider that first.

```{r layout="l-body-outset"}
filter(x, alg == "cuhre_0") %>%
  ggplot(aes(1/epsrel, errorest, shape = version)) +
  geom_point() +
  scale_shape_manual(values = c(1,4)) +
  scale_x_log10() +
  scale_y_log10() +
  facet_wrap(vars(integrand), scales = "free", ncol=1) +
  labs(x = "Inverse of error tolerance",
       y = "Fractional error estimate",
       title="cuhre_0")
```

```{r layout="l-body-outset"}
filter(x, alg == "cuhre_1") %>%
  ggplot(aes(1/epsrel, errorest, shape = version)) +
  geom_point() +
  scale_shape_manual(values = c(1,4)) +
  scale_x_log10() +
  scale_y_log10() +
  facet_wrap(vars(integrand), scales = "free", ncol=1) +
  labs(x = "Inverse of error tolerance",
       y = "Fractional error estimate",
       title="cuhre_1")
```

```{r layout="l-body-outset"}
filter(x, alg == "vegas") %>%
  ggplot(aes(1/epsrel, errorest, shape = version)) +
  geom_point() +
  scale_shape_manual(values = c(1,4)) +
  scale_x_log10() +
  scale_y_log10() +
  facet_wrap(vars(integrand), scales = "free", ncol=1) +
  labs(x = "Inverse of error tolerance",
       y = "Fractional error estimate",
       title = "vegas")
```

The difference in the plots for `fun6` is a result of an upgrading of the
performance measurement harness, which now uses more closely-spaced error 
tolerances and which has a less extreme upper end to the error tolerance used.

In no case does the new algorithm seem to differ in its results from the old.

## Comparison of integral estimates

```{r layout="l-body-outset"}
filter(x, alg == "cuhre_0") %>%
  ggplot(aes(1/epsrel, value, shape = version)) +
  geom_point() +
  scale_shape_manual(values = c(1,4)) +
  scale_x_log10() +
  facet_wrap(vars(integrand), scales = "free_x", ncol=1) +
  labs(x = "Inverse of error tolerance",
       y = "Integral estimate",
       title="cuhre_0")
```

```{r layout="l-body-outset"}
filter(x, alg == "cuhre_1") %>%
  ggplot(aes(1/epsrel, value, shape = version)) +
  geom_point() +
  scale_shape_manual(values = c(1,4)) +
  scale_x_log10() +
  facet_wrap(vars(integrand), scales = "free_x", ncol=1) +
  labs(x = "Inverse of error tolerance",
       y = "Integral estimate",
       title="cuhre_1")
```

```{r layout="l-body-outset"}
filter(x, alg == "vegas") %>%
  ggplot(aes(1/epsrel, value, shape = version)) +
  geom_point() +
  scale_shape_manual(values = c(1,4)) +
  scale_x_log10() +
  facet_wrap(vars(integrand), scales = "free_x", ncol=1) +
  labs(x = "Inverse of error tolerance",
       y = "Integral estimate",
       title = "vegas")
```

For the integral estimates, there is no observable difference between the
old and new releases.

## Number of function calls

```{r layout="l-body-outset"}
filter(x, alg == "cuhre_0") %>%
  ggplot(aes(1/epsrel, neval, shape = version)) +
  geom_point() +
  scale_shape_manual(values = c(1,4)) +
  scale_x_log10() +
  scale_y_log10() +
  facet_wrap(vars(integrand), scales = "free", ncol=1) +
  labs(x = "Inverse of error tolerance",
       y = "Number of function evaluations",
       title="cuhre_0")
```

```{r layout="l-body-outset"}
filter(x, alg == "cuhre_1") %>%
  ggplot(aes(1/epsrel, neval, shape = version)) +
  geom_point() +
  scale_shape_manual(values = c(1,4)) +
  scale_x_log10() +
  scale_y_log10() +
  facet_wrap(vars(integrand), scales = "free", ncol=1) +
  labs(x = "Inverse of error tolerance",
       y = "Number of function evaluations",
       title="cuhre_1")
```

```{r layout="l-body-outset"}
filter(x, alg == "vegas") %>%
  ggplot(aes(1/epsrel, neval, shape = version)) +
  geom_point() +
  scale_shape_manual(values = c(1,4)) +
  scale_x_log10() +
  scale_y_log10() +
  facet_wrap(vars(integrand), scales = "free", ncol=1) +
  labs(x = "Inverse of error tolerance",
       y = "Number of function evaluations",
       title = "vegas")
```

There is no observable difference in the number of function calls required by
any of the algorithms.

## Timing

```{r layout="l-body-outset"}
filter(x, alg == "cuhre_0") %>%
  ggplot(aes(1/epsrel, time, shape = version)) +
  geom_point() +
  scale_shape_manual(values = c(1,4)) +
  scale_x_log10() +
  scale_y_log10() +
  facet_wrap(vars(integrand), scales = "free", ncol=1) +
  labs(x = "Inverse of error tolerance",
       y = "Calculation time",
       title="cuhre_0")
```

```{r layout="l-body-outset"}
filter(x, alg == "cuhre_1") %>%
  ggplot(aes(1/epsrel, time, shape = version)) +
  geom_point() +
  scale_shape_manual(values = c(1,4)) +
  scale_x_log10() +
  scale_y_log10() +
  facet_wrap(vars(integrand), scales = "free", ncol=1) +
  labs(x = "Inverse of error tolerance",
       y = "Calculation time",
       title="cuhre_1")
```

```{r layout="l-body-outset"}
filter(x, alg == "vegas") %>%
  ggplot(aes(1/epsrel, time, shape = version)) +
  geom_point() +
  scale_shape_manual(values = c(1,4)) +
  scale_x_log10() +
  scale_y_log10() +
  facet_wrap(vars(integrand), scales = "free", ncol=1) +
  labs(x = "Inverse of error tolerance",
       y = "Calculation time",
       title = "vegas")
```

There seems to be a small speed improvement in the newer algorithm version.
However, it is possible that this speed difference is due to the newer
compiler version used to build the newer CUBA code.