---
title: "Where serial CUHRE evaluates functions"
description: |
  A visualization of where CUHRE evaluates functions for a 2-dimensional integrand.
author:
  - name: Marc Paterno
    url: https://github.com/marcpaterno
date: 06-10-2020
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

This document shows where the serial CUHRE algorithm evaluates the function it is integrating.
Unlike a Monte Carlo algorithm (such as VEGAS), CUHRE evaluates the function at a set of deterministically chosen points.

The function used as the integrand in this example is:

$$ 2 x^2 + 2 y - y^2.$$
We have used both VEGAS and CUHRE to integrate this function, using a fractional error tolerance of 0.01.
The large fractional error was chosen to limit the number of function evaluations.
This function is sufficiently simple that CUHRE uses only two regions.
The plots show where VEGAS and CUHRE each evaluate the function to achieve this level of accuracy.

```{r echo = FALSE, message = FALSE}
d <- bind_rows(cuhre = read_tsv("cuhre_points.txt.bz2"),
               vegas = read_tsv("vegas_points.txt.bz2"),
               .id = "alg")
```

VEGAS required `r nrow(filter(d, alg == "vegas"))` function evaluations, and CUHRE required `r nrow(filter(d, alg == "cuhre"))`.

```{r, echo = FALSE}
ggplot(d, aes(x,y)) +
  geom_point(size = 0.5, alpha = 0.3) +
  facet_wrap(vars(alg))
```
