---
title: "A first look at parallel CUHRE subregions"
description: |
  A first analysis the shapes of subregions generated by the parallel CUHRE algorithm.
  We are looking to determine whether parallel CUHRE is prone to producing very
  enlongated regions.
author:
  - name: Marc Paterno
    url: https://github.com/marcpaterno
date: 07-06-2020
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## The integrand

The integrand chosen is:
$$ | \cos(4 v +5 w + 6 x +7 y + 8 z)/k |$$
with $k = 0s .6371054$.
For this integrand, the normalization is approximate.
The true value of the integrand is close to, but not exactly, 1.

## The data recorded

```{r get_data}
# Download the data if it does not already exist locally.
if (!file.exists("df5.fst")) {
  filename <- "genz_1abs_5d_regions.csv.xz"
  downloadstatus <- 
    download.file(sprintf("https://zenodo.org/record/4319038/files/%s", filename),
                  filename)
  stopifnot(downloadstatus == 0)
  tmp <- readr::read_csv(filename)
  unlink(filename)
  tmp <- as_tibble(tmp)
  fst::write_fst(tmp, "df5.fst")
  rm(tmp)
}
```

```{r make_df5}
df5 <- fst::read_fst("df5.fst") %>%
  mutate(vol = dim0 * dim1 * dim2 * dim3 * dim4)
```

```{r summarize_df5}
summary(df5)
```

## Variations with `div`

`div` is the number of divisions this volume has been subject to (since the start of Phase 2).

For this integrand, the estimated value of the integral over the subregion decreases
with the number of divisions:

```{r}
ggplot(df5, aes(div, value, group = div)) +
  geom_boxplot() +
  scale_y_log10() +
  labs(x="Number of divisions for the volume", y="value of integral in volume")
```

## How "stretched" are the subregions?

We define the quantity *stretch* to be the ratio of the maximum subregion extent to the minimum subregion extent.
A hypercube would have a *stretch* of 1.

```{r add_stretch}
df5 <- mutate(df5, mind = pmin(dim0, dim1, dim2, dim3, dim4),
            maxd = pmax(dim0, dim1, dim2, dim3, dim4),
            stretch = maxd/mind)
```

```{r}
ggplot(df5, aes(stretch)) +
  geom_histogram(bins = 50, binwidth = 1, center = 0)
```

No region from this application of the algorithm has a stretch of more than 4.
At least for this integrand, there is no sign that the algorithm is making very
enlongated regions.
