---
title: "Comparing serial VEGAS and CUHRE on a (partially) non-separable integrand"
description: |
  This document shows a comparison of the speed of the serial VEGAS and CUHRE
  algorithms, as implemented in the CUBA (http://www.feynarts.de/cuba/) library,
  and wrapped by `cubacpp` (https://bitbucket.org/mpaterno/cubacpp), on a
  non-separable integrand.
  
  Note this was updated on 2020-12-13.
author:
  - name: Marc Paterno
    url: https://github.com/marcpaterno
date: 06-24-2020
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rmarkdown)
library(tidyverse)
```

## The algorithms

`vegas` is the Vegas algorithm of Lepage, as implemented in the CUBA library.
This version uses quasi-random (low-descrepency sequence) numbers, rather than pseudo-random numbers.
`cuhre_0` is the serial CUHRE algorithm, as implemented in the CUBA library, using `flag` = 0.
This version uses all the volumes produced by the algorithm for determining the final result.
`cuhre_1` is the serial CUHRE algorithm, as implemented in the CUBA library, using `flag` = 4.
This version uses only the final set of volumes produced by the algorithm for determining the final result.

## The integrand

The integrand chosen is:
$$ k (u \, v + \frac{x \, y \, w^y}{1+u} +z^2) $$
with
$$k = \frac{12}{7 - 6 (\ln(2))^2 + \ln(64)}$$.

The correct value of the integral is 1.

## Testing environment

These tests were run on a MacBook Pro laptop.

```{r child = '../../mac130389-description.Rmd'}
```

## Description of the dataframe

```{r child = '../../dataframe-description.Rmd'}
```

```{r read_data, echo = FALSE, message = FALSE}
d <- read_tsv("fun6_vegas_cuhre_comparison.tsv.xz") %>%
  mutate(r = errorest/(epsrel*value),
         alg = factor(alg, levels =c("vegas", "cuhre_0", "cuhre_1") )) %>%
  arrange(desc(epsrel), alg)

paged_table(d)
```

## Analysis

As one measure of the innate "efficiency" of each algorithm,
we consider how many function evaluations are required
to obtain a given fractional error tolerance.

```{r, echo = FALSE}
ggplot(d, aes(1/epsrel, neval, color = alg)) +
  geom_point() +
  scale_x_log10() +
  scale_y_log10() +
  labs(x = "Inverse of fractional error tolerance", y = "Number of function evaluations required")
```

For this integrand, VEGAS always requires more function calls than does CUHRE.
CUHRE with flag=0 requires fewer function evaluations than it does with flag=4.

### Calculation time as a function of required error tolerance

```{r, echo = FALSE}
ggplot(d, aes(1/epsrel, time, color=alg)) +
  geom_point() +
  scale_x_log10() +
  scale_y_log10() +
  labs(x="Inverse of error tolerance", y="Calculation time (ms)")

```

### Error estimate and function evaluations

```{r echo = FALSE, message = FALSE}
d %>%
ggplot(aes(neval, errorest, color = alg)) +
  geom_smooth(method = "lm") +
  geom_point() +
  scale_x_log10() +
  scale_y_log10() +
  labs(x = "Number of integrand evaluations", "Fractional error estimate")
```

### Reliability of error estimate

If we look only at the calculations that converged, how accurate are the error
estimates?

```{r, echo=FALSE}
filter(d, !is.na(value)) %>%
  ggplot(aes(1/epsrel, abs(error)/errorest, color = alg)) +
  scale_y_log10(labels = scales::comma) +
  scale_x_log10() +
  geom_point() +
  labs(x = "Inverse of error tolerance",
       y = "Ratio of actual error to estimated error")
```

If the error estimate is reliable, then the ratio should not exceed 1.
At the most stringent (fractional) error tolerances, the error estimate from
the CUHRE algorithm with flag=0 are not reliable.
That is, the algorithm is claiming to have converged when it actually has not
converged.

### Error estimate as a function of running time

Users do not really care about the number of function evaluations done;
they care about the time the algorithm takes to converge for a given
error tolerance, and that the algorithm has actually converged to that tolerance.

```{r, echo = FALSE}
filter(d, !is.na(value)) %>%
  ggplot(aes(1/epsrel, time, color = alg)) +
  scale_x_log10() +
  scale_y_log10(labels = scales::comma) +
  geom_point() +
  labs(x = "Inverse of error tolerance",
       y = "Calculation time (ms)")
```

For this integrand, the CUHRE algorithm is much faster than VEGAS.
The CUHRE algorithm with flag = 0 is faster for large error tolerances,
but as we have seen above, the error estimate (and thus the convergence)
becomes unreliable.
Since in real problems the user would *not* know the value of the integral
(or they would not be using a numerical approximation method!),
it seems prudent to use the algorithm with flag = 4 -- at least for this
integrand.
