---
title: "The Genz function 1 in 8d"
description: |
  This document shows a performance comparison between the serial
  and parallel implementations of the CUHRE algorithm for a
  non-positive-definite integrand.
author:
  - name: Marc Paterno
    url: https://github.com/marcpaterno
date: 06-08-2020
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## The integrand

The integrand chosen is:
$$ \cos(s + 2 t + 3 u +4 w +5 w + 6 x +7 y + 8 z) / k$$
with
$$ k = (1/315)  \sin(1)  \sin(3/2) \sin(2) \sin(5/2) \sin(3) 
       \sin(7/2) \sin(4) (\sin(37/2) - \sin(35/2)). $$

$k$ is approximately equal to
`r format((1/315) * sin(1) * sin(3/2) * sin(2) *sin(5/2) * sin(3) * sin(7/2) * sin(4) * (sin(37/2) - sin(35/2)), digits=15)`.

## Testing environment

These tests were run on `ibmpower9.fnal.gov`.

```{r child = '../../power9_description.Rmd'}
```

## Description of the dataframe

```{r child = '../../dataframe-description.Rmd'}
```

```{r read_data, echo = FALSE, message = FALSE}
# d <- bind_rows(read_tsv("genz_1_8d.out.bz2"),
#                read_tsv("cuda_genz_1_8d.out.bz2"))
d <- read_tsv("genz_1_8d.out.bz2") %>%
  mutate(r = errorest/(epsrel*value)) %>%
  arrange(alg, desc(epsrel))

knitr::kable(d) #, digits = 8)
```

## Analysis

We do not yet have results from the parallel algorithm.
These results are from the serial algorithm.

Because the range of fractional error tolerance values is large, and the range of times is very large, we use log scales for our plot.
The line is a fitted linear model; the fit is clearly poor.

```{r, echo = FALSE, message = FALSE}
ggplot(d, aes(1/epsrel, time)) +
  geom_point() +
  scale_x_log10() +
  scale_y_log10() +
  geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(vars(alg))
```

The number of regions needed to obtain a given fractional error tolerance seems to be related in a power law
to the reciprocal of the fractional error tolerance.

```{r, echo = FALSE, message = FALSE}
ggplot(d, aes(1/epsrel, nregions)) +
  geom_point() +
  scale_x_log10() +
  scale_y_log10() +
  geom_smooth(method = "lm") +
  facet_wrap(vars(alg))
```
